Kronos \& Krios: Evidence for accretion of a massive, rocky planetary system in a comoving pair of G dwarfs

### 2017-06-15 paper/ms.tex
Many typos and dumb mistakes

Here, we report and discuss the discovery of a comoving pair of G stars,
HD~240430 and HD~240429, with unusual chemical abundance differences
that strongly suggest accretion of planetary material by the two stars.
Throughout the \documentname, we nickname the two stars \bizarreone\
(HD~240430) and \sunanalog\ (HD~240429).
We use the following convention for chemical abundances of stars: \elemH{X} is
the ratio of number density of element \elem{X} to \elem{H} relative to the
solar value.
The absolute abundance of \elem{X} is $A(\elem{X}) = 12 + \log_{10}
(n_\elem{X}/n_\elem{H})$ where $n_\elem{X}$ is the number density of element
\elem{X}.
In \sectionname~\ref{sec:data} we present the astrometric and spectroscopic data
about the two stars relevant to the present discussion.
In \sectionname~\ref{sec:discussion} we discuss possible interpretations of
the abundance difference between the pair.
We summarize our discussions in \sectionname~\ref{sec:summary}.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{deltaXH_elem_violins.pdf}
  \caption{Abundance difference in this pair and other twin-like
    ($\Delta T_\mathrm{eff}\lesssim 100$~K) wide binaries in
    \citealt{2016ApJS..225...32B}.
    The differences in other pairs are small ($<0.05$~dex)
    for all elements except \elem{N} and \elem{O} which are the most
    uncertain, making the difference of $\approx 0.2$~dex seen in
    \bizarreone-\sunanalog\ rare.
    Additionally, we show the distribution of abundance differences
    between field stars with similar metallicity difference
    ($\Delta[\elem{Fe}/\elem{H}] \approx 0.2$)
    as violins with medians indicated by black line segments.
    These are random pairings of single stars in
    in \citealt{2016ApJS..225...32B} at two metallicity bins,
    $-0.025 < \feh < 0.025$ (160 stars) and $1.975 > \feh > 2.025$ (137 stars),
    similar to \bizarreone\ and \sunanalog.
    The difference is always taken to be
    $\mathrm{higher}\,\feh - \mathrm{lower}\,\feh$.
    Thus, the narrower range of in $\Delta\feh$ is by construction.
    Random pairings of disk stars with similar $\Delta\feh$ usually show
    similar enhancement in all other elements
    unlike the pattern seen in \bizarreone-\sunanalog\ pair.
  }
  \label{fig:deltaXH}
\end{figure}

@@

In fact, the surface lithium abundance is the only indication that points to
younger ages.
If the two stars are indeed only several hundred Myrs old at most,
they are expected to be part of a larger comoving group of stars.
However, as we mention above, there is no evidence in our search of comoving pairs
using \tgas\ that the two stars belong to a larger group of young stars.
Very young stars are often show signs of activity such as
X-ray emission from magnetic activity, emission lines, or infrared excess due to
circumstellar disks (\todo{needs refernces}).
We have compiled GALEX, Tycho-2, 2MASS, and WISE photometry for these stars,
and found no evidence for indications of activity in their spectral energy
distributions.
The low $v\sin(i)$ values (\tablename~\ref{tab:kk}) also argue against very
young ages that would be inferred from the surface lithium abundances.
Finally, we computed the Galactic orbit of the pair using the median of the
posterior sample over the phase space coordinates of \sunanalog, in a Milky
Way-like gravitational potential (similar to \texttt{MWPotential2014} from
\citealt{Bovy:2015}) using \project{Gala} (\citealt{gala}).
The pair's fiducial orbit has a vertical action larger than the Sun, favoring
an older age (\todo{needs references for vertical action as age indicator}).
We therefore conclude that the two stars are most likely coeval, $\sim 4$~Gyr
old main sequence stars, and that their unusually high \elem{Li} abundance
requires an alternative explanation.

@@

Another possibility that two coeval stars may end up with different surface
abundances is accretion of planetary material after birth.
Instabilities may develop in a multi-planet system due to its chaotic nature
which may lead to planet engulfment or ejection by the host star
(\todo{needs references}).
%TODO: kozai?
Indeed, it is an important goal of many exoplanet studies
to detect chemical signatures of planet formation or accretion,
distinguish them from Galactic chemical evolution, and
connect them to theories of evolution of planetary systems.
One approach that is free from confusion with Galactic chemical evolution
is to compare two almost identical stars in a wide binary system.
Assuming that the component stars were born together with identical
initial composition, we may see a difference in their surface abundances
if the two stars then accreted different amounts of accretion
of planetary material.
The resulting abundance difference may depend on the condensation
temperatures of elements in the protoplanetary disks from which the accreted
planets formed, as their compositions depend on the radial temperature gradient
in the disk.
%TODO: cautions on condensation temperature
% 1. it is just a proxy
% 2. it is **equlibrium** condensation temperature

In Figure~\ref{fig:relabun_tcrank}, we show the abundace difference
between \bizarreone\ and \sunanalog\ ordered by the rank of \Tcondens\
of each element.
The equilibrium condensation temperatures for the composition of solar system
are taken from \citealt{2003ApJ...591.1220L} (Table~8).
The difference seen in \bizarreone-\sunanalog\ is
compared to HD~20781/2, XO-2N/S, WASP-94A/B in \figname~\ref{fig:relabun_tcrank}.
The metallicity difference of $\approx 0.2$~dex observed in this pair
is larger than the differences seen in any other pairs studied so far.
Strikingly, the five under-enhanced elements in \bizarreone\
relative to \sunanalog\ are the five most volatile in all elements measured.
The difference in \elem{Mn} ($\Tcondens = 1158$~K) and
\elem{Cr} ($\Tcondens = 1296$~K) suggests a break in $\Tcondens \approx 1200$~K.
This $\Tcondens$-dependent trend of $\Delta\elemH{X}$,
combined with the enhanced $\elem{Li}$ abundance ($A(\elem{Li}) = 2.75$),
strongly suggests that accretion of rocky material has occured in \bizarreone.

@@

Figure~\ref{fig:toycalc} shows the expected change of surface abundances of
metals in a Sun-like star after $20~\mearth$ of material with composition of
bulk Earth is added.
A volatility trend that more volatile (low \Tcondens) elements are more
depleted in the Earth relative to CI or other carbonaceous chondrites
has long been known (\citealt{mcdonough2001composition}).
This trend is presumed to be closely related to the formation of terrestrial
planets and, in particular to the radial temperature gradient in a
protoplanetary disk.
The trend resulting from adding $20~\mearth$ of bulk Earth
provides an overall good match to the observed $\Delta\elemH{X}$,
suggeting that the refractory-enhanced star, \bizarreone\,
accreted $20~\mearth$ more of rocky planetary material than \sunanalog.

@@

%TODO: augment this paragraph
% - more geochemical study references
We stress that while the calculation carried out is useful in
order-of-magnitude sense, further investigation of each of the simplifying
assumptions made is warranted.
In addition, the composition of bulk Earth has some uncertainties.
For example, the reported bulk Earth concentration of the siderphile element
\elem{Mn}, varies from $800$ to $1700$~ppm (\citealt{1998psc..book.....L}).
While the latter value from \citealt{mcdonough2001composition} has been used in
our calculation, the former value would bring the observed $\Delta\elemH{Mn}$
to an even closer agreement.
Given these limitations, the level of agreement for $\Delta\elemH{X}$ {\it and}
\elem{Li} for \bizarreone\ is remarkable.

>>>>

Here, we report and discuss the discovery of a comoving pair of G stars,
HD~240430 and HD~240429, with unusual chemical abundance differences
that strongly suggest accretion of planetary material by the two stars.
Throughout the \documentname, we nickname the two stars \bizarreone\
(HD~240430) and \sunanalog\ (HD~240429).
We use the following convention for chemical abundances of stars: \elemH{X} is
the log ratio of number density of element \elem{X} to \elem{H} relative to the
solar value.
The absolute abundance of \elem{X} is $A(\elem{X}) = 12 + \log_{10}
(n_\elem{X}/n_\elem{H})$ where $n_\elem{X}$ is the number density of element
\elem{X}.
In \sectionname~\ref{sec:data} we present the astrometric and spectroscopic data
about the two stars relevant to the present discussion.
In \sectionname~\ref{sec:discussion} we discuss possible interpretations of
the abundance difference between the pair.
We summarize our discussions in \sectionname~\ref{sec:summary}.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{deltaXH_elem_violins.pdf}
  \caption{Abundance difference in this pair and other twin-like
    ($\Delta T_\mathrm{eff}\lesssim 100$~K) wide binaries in
    \citealt{2016ApJS..225...32B}.
    The differences in other pairs are small ($<0.05$~dex)
    for all elements except \elem{N} and \elem{O} which are the most
    uncertain, making the difference of $\approx 0.2$~dex seen in
    \bizarreone-\sunanalog\ rare.
    Additionally, we show the distribution of abundance differences
    between field stars with similar metallicity difference
    ($\Delta[\elem{Fe}/\elem{H}] \approx 0.2$)
    as violins with medians indicated by black line segments.
    These are random pairings of single stars in
    in \citealt{2016ApJS..225...32B} at two metallicity bins,
    $-0.025 < \feh < 0.025$ (160 stars) and $0.175 > \feh > 0.225$ (137 stars),
    similar to \bizarreone\ and \sunanalog.
    The difference is always taken to be
    $\mathrm{higher}\,\feh - \mathrm{lower}\,\feh$.
    Thus, the narrower range of in $\Delta\feh$ is by construction.
    Random pairings of disk stars with similar $\Delta\feh$ usually show
    similar enhancement in all other elements
    unlike the pattern seen in \bizarreone-\sunanalog\ pair.
  }
  \label{fig:deltaXH}
\end{figure}

@@

In fact, the surface lithium abundance is the only indication that points to
younger ages.
If the two stars are indeed only several hundred Myrs old at most,
they are expected to be part of a larger comoving group of stars.
However, as we mention above, there is no evidence in our search of comoving pairs
using \tgas\ that the two stars belong to a larger group of young stars.
Very young stars often show signs of activity such as
X-ray emission from magnetic activity, emission lines, or infrared excess due to
circumstellar disks (\todo{needs refernces}).
We have compiled GALEX, Tycho-2, 2MASS, and WISE photometry for these stars,
and found no evidence for indications of activity in their spectral energy
distributions.
The low $v\sin(i)$ values (\tablename~\ref{tab:kk}) also argue against very
young ages that would be inferred from the surface lithium abundances.
Finally, we computed the Galactic orbit of the pair using the median of the
posterior sample over the phase space coordinates of \sunanalog, in a Milky
Way-like gravitational potential (similar to \texttt{MWPotential2014} from
\citealt{Bovy:2015}) using \project{Gala} (\citealt{gala}).
The pair's fiducial orbit has a vertical action larger than the Sun, favoring
an older age (\todo{needs references for vertical action as age indicator}).
We therefore conclude that the two stars are most likely coeval, $\sim 4$~Gyr
old main sequence stars, and that their unusually high \elem{Li} abundance
requires an alternative explanation.

@@

Another possibility that two coeval stars may end up with different surface
abundances is accretion of planetary material after birth.
Instabilities may develop in a multi-planet system due to its chaotic nature
which may lead to planet engulfment or ejection by the host star
(\todo{needs references}).
%TODO: kozai?
Indeed, it is an important goal of many exoplanet studies
to detect chemical signatures of planet formation or accretion,
distinguish them from Galactic chemical evolution, and
connect them to theories of evolution of planetary systems.
One approach that is free from confusion with Galactic chemical evolution
is to compare two almost identical stars in a wide binary system.
Assuming that the component stars were born together with identical
initial composition, we may see a difference in their surface abundances
if the two stars then accreted different amounts of planetary material.
The resulting abundance difference may depend on the condensation
temperatures of elements in the protoplanetary disks from which the accreted
planets formed, as their compositions depend on the radial temperature gradient
in the disk.
%TODO: cautions on condensation temperature
% 1. it is just a proxy
% 2. it is **equlibrium** condensation temperature

In Figure~\ref{fig:relabun_tcrank}, we show the abundance difference
between \bizarreone\ and \sunanalog\ ordered by the rank of \Tcondens\
of each element.
The equilibrium condensation temperatures for the composition of solar system
are taken from \citealt{2003ApJ...591.1220L} (Table~8).
The difference seen in \bizarreone-\sunanalog\ is
compared to HD~20781/2, XO-2N/S, WASP-94A/B in \figname~\ref{fig:relabun_tcrank}.
The metallicity difference of $\approx 0.2$~dex observed in this pair
is larger than the differences seen in any other pairs studied so far.
Strikingly, the five under-enhanced elements in \bizarreone\
relative to \sunanalog\ are the five most volatile in all elements measured.
The difference in \elem{Mn} ($\Tcondens = 1158$~K) and
\elem{Cr} ($\Tcondens = 1296$~K) suggests a break in $\Tcondens \approx 1200$~K.
This $\Tcondens$-dependent trend of $\Delta\elemH{X}$,
combined with the enhanced $\elem{Li}$ abundance ($A(\elem{Li}) = 2.75$),
strongly suggests that accretion of rocky material has occured in \bizarreone.

@@

Figure~\ref{fig:toycalc} shows the expected change of surface abundances of
metals in a Sun-like star after $20~\mearth$ of material with composition of
bulk Earth is added.
A volatility trend that more volatile (low \Tcondens) elements are more
depleted in the Earth relative to CI or other carbonaceous chondrites
has long been known (\citealt{mcdonough2001composition}).
This trend is presumed to be closely related to the formation of terrestrial
planets and, in particular to the radial temperature gradient in a
protoplanetary disk.
The trend resulting from adding $20~\mearth$ of bulk Earth
provides an overall good match to the observed $\Delta\elemH{X}$,
suggesting that the refractory-enhanced star, \bizarreone\,
accreted $20~\mearth$ more of rocky planetary material than \sunanalog.

@@

%TODO: augment this paragraph
% - more geochemical study references
We stress that while the calculation carried out is useful in
an order-of-magnitude sense, further investigation of each of the simplifying
assumptions made is warranted.
In addition, the composition of bulk Earth has some uncertainties.
For example, the reported bulk Earth concentration of the siderphile element
\elem{Mn}, varies from $800$ to $1700$~ppm (\citealt{1998psc..book.....L}).
While the latter value from \citealt{mcdonough2001composition} has been used in
our calculation, the former value would bring the observed $\Delta\elemH{Mn}$
to an even closer agreement.
Given these limitations, the level of agreement for $\Delta\elemH{X}$ {\it and}
\elem{Li} for \bizarreone\ is remarkable.
